Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2019-01-16 scagnostics.Rmd
start on the information indexes

* **Information**:  The index functions [@Reshef1518] measure nonlinear association using a nonparametric computation of information. It is strictly a 2D measure, which observations are binned and the relative count of the bin is used to compute entropy. are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$. With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as
\begin{equation}
M(D)_{x,y} = \frac{I^{\ast} (D, x, y)}{\log \min{x,y}},
\end{equation}
where the denominator is chosen such that each entry of $M$ will take values between [0,1].
We can now define the following metrics:
    + Maximum Information Coefficient (MIC):
    \begin{equation}
    MIC(D) = \max_{xy<B(n)}\{M(D)_{x,y}\},
    \label{eq:MIC}
    \end{equation}
    where by default $B(n) = n^{0.6}$
    + Total Information Coefficient (TIC), see @JMLRv1715308:
    \begin{equation}
    TIC(D) = \sum_{xy<B(n)} M(D)_{x,y}
    \end{equation}
  While $MIC$ is normalized to fall between $[0,1]$, the upper bound of $TIC$ depends on the sample size via the bounding function $B(n)$. $TIC$ was designed to be more stable test against null hypothesis of independence, using the full information of the characteristic matrix rather than just the maximum, and as such is expected to be better suited as a PPI.

>>>>

* **Information**:  The index functions [@Reshef1518] nonparametricly measure nonlinear association by finding an optimal grid to separate empty cells from cells containing observations. It is strictly a 2D measure, where the relative bin counts are used to compute mutual information, $\bm{I} = \sum_{by_1}\sum_{by_2} p(by_1, by_2) log(p(by_1, by_2)/(p(by_1)p{by_2}))$, where $by_1, by_2$ are binned values of the projected data, and $p(by_1, by_2)$ is the relative bin count in each cell, and $p(by_1), p(by_2)$ are the row and column relative counts.

are based on the mutual information $I$ when considering the distribution induced by the data points $D \subset \mathbb{R}^2$ on an $x-$by$y$ grid $G$. With $I^{\ast} (D, x, y) = \max I (D|_G)$ the characteristic matrix $M$ is defined as
\begin{equation}
M(D)_{x,y} = \frac{I^{\ast} (D, x, y)}{\log \min{x,y}},
\end{equation}
where the denominator is chosen such that each entry of $M$ will take values between [0,1].
We can now define the following metrics:
    + Maximum Information Coefficient (MIC):
    \begin{equation}
    MIC(D) = \max_{xy<B(n)}\{M(D)_{x,y}\},
    \label{eq:MIC}
    \end{equation}
    where by default $B(n) = n^{0.6}$
    + Total Information Coefficient (TIC), see @JMLRv1715308:
    \begin{equation}
    TIC(D) = \sum_{xy<B(n)} M(D)_{x,y}
    \end{equation}
  While $MIC$ is normalized to fall between $[0,1]$, the upper bound of $TIC$ depends on the sample size via the bounding function $B(n)$. $TIC$ was designed to be more stable test against null hypothesis of independence, using the full information of the characteristic matrix rather than just the maximum, and as such is expected to be better suited as a PPI.
